---
title: "Additional Color Palettes"
output: 
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Additional Color Palettes}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.asp = 2/3,
  out.width = "75%",
  dev = "ragg_png"
)
```

```{r setup, include=FALSE}
library(ggplot2)
library(crthemes)
```

*Updated `r format(Sys.time(), "%b %d, %Y")`*
\
\
`crthemes` includes several qualitative, discrete palettes, but does not include any ordered discrete palettes or continuous palettes. The [`paletteer` package](https://emilhvitfeldt.github.io/paletteer/) collects a large variety of discrete and continuous color palettes and provides wrapper functions to access them. 

[`show_col()`](https://scales.r-lib.org/reference/show_col.html) from the [`scales` package](https://scales.r-lib.org/index.html) can be used to preview colors from other packages (and is the basis for the `pal_preview()` function in this package).

## Using `paletteer`

`paletteer_c()` and `paletteer_d()` access continuous and discrete palettes in the `paletteer` package, respectively. The arguments are:

* `palette`: the specific palette to be used, must be provided as a string in the form `"packagename::palettename"`
* `n`: the number of colors to return; required for continuous palettes, optional for discrete palettes
* `direction`: 1 (default) for the standard direction, -1 for reversed

`palettes_c_names` and `palettes_d_names` print the packages and palettes for the included continuous and discrete palettes, respectively. The included palettes are also listed in [this repository](https://github.com/EmilHvitfeldt/r-color-palettes).

To use more colors than are included in a palette, use the `colour_ramp()` function from the scales package to interpolate the colors. `colour_ramp()` returns a function that will take input of a numeric sequence and return a character vector containing the palette colors. 

```{r paletteer, fig.width=7.5, out.width="60%"}
library(scales)
library(paletteer)

pal <- paletteer_d("rcartocolor::DarkMint")
show_col(pal)
# To interpolate between colors in a palette, use colour_ramp()
ramp <- colour_ramp(pal)
# ramp is a function that takes a numeric vector and returns a palette
pal25 <- ramp(seq(0, 1, length.out = 25))
show_col(pal25)
# Paletteer also has continuous palettes
pal <- paletteer_c("pals::ocean.deep", n = 25)
show_col(pal)
```

### Scale Functions

`paletteer` includes `scale_*_paletteer_c()` and `scale_*_paletteer_d()` functions to apply the included palettes to ggplots. Like with the other `paletteer` functions, the palette needs to be provided as a string in the form `packagename::palettename`

#### Continuous

```{r scales-paletteer-c}
set.seed(503)
x <- rnorm(200)
df <- data.frame(x = x, y = x + rnorm(200, sd = 1.5), 
                 z = x + rnorm(200, mean = 5, sd = 0.5))
p <- ggplot(df, aes(x, y, color = z)) +
  geom_point(alpha = 0.6) +
  labs(x = "X Axis", y = "Y Axis", color = NULL, title = "Continuous") +
  theme_cr(cairo = TRUE, font = "")

# Continuous palette
p + scale_color_paletteer_c("pals::ocean.matter", direction = -1)
```

#### Discrete

```{r scales-paletteer-d}
library(dplyr)

df <- df %>% mutate(bin = ntile(.data$z, 7))
p_d <- ggplot(df, aes(x, y, color = as.factor(bin))) +
  geom_point(alpha = 0.6) +
  labs(x = "X Axis", y = "Y Axis", color = NULL, title = "Discrete") +
  theme_cr(cairo = TRUE, font = "")

# Discrete palette
p_d + scale_color_paletteer_d("rcartocolor::Sunset")
```

#### Interpolating

If the number of colors needed is not the same as the number of colors in the palette, the `colour_ramp()` function can be used to interpolate the palette. 

```{r scales-interpolate}
df <- df %>% mutate(bin = ntile(.data$z, 5))

# Interpolate palette to 5 colors
ramp <- colour_ramp(paletteer_d("rcartocolor::Sunset"))
pal <- ramp(seq(0, 1, length.out = 5))

p_i <- ggplot(df, aes(x, y, color = as.factor(bin))) +
  geom_point(alpha = 0.6) +
  labs(x = "X Axis", y = "Y Axis", color = NULL, title = "Interpolated") +
  theme_cr(cairo = TRUE, font = "")

# Interpolated palette; use scale_color_manual because the colors are in a variable
p_i + scale_color_manual(values = pal)
```

## Palette Options

Below are some options for sequential and diverging color palettes from the `paletteer`, `RColorBrewer`, and `viridis` packages. 

The `RColorBrewer` package is included in `ggplot2` by default through:

* `scale_color_brewer()` (discrete)
* `scale_color_distiller()` (continuous)
* `scale_color_fermenter()` (binned)

The `viridis` package is included in `ggplot2` by default through: 

* `scale_color_viridis_d()` (discrete)
* `scale_color_viridis_c()` (continuous)
* `scale_color_viridis_b()` (binned)

For more colors, see the full [list of available palettes](https://github.com/EmilHvitfeldt/r-color-palettes).

### Sequential

```{r paletteer-seq, fig.width=7.5, out.width="60%"}
# Discrete, "BurgYl" from "rcartocolor
pal <- paletteer_d("rcartocolor::BurgYl")
show_col(pal, labels = FALSE)
# Discrete, "DarkMint" from "rcartocolor
pal <- paletteer_d("rcartocolor::DarkMint")
show_col(pal, labels = FALSE)
# Discrete, "RdPu" from "RColorBrewer
pal <- paletteer_d("RColorBrewer::RdPu")
show_col(pal, labels = FALSE)
# Discrete, "PuBuGn" from "RColorBrewer
pal <- paletteer_d("RColorBrewer::PuBuGn")
show_col(pal, labels = FALSE)
# Continuous, "magma" from "viridis"
pal <- paletteer_c("viridis::magma",16)
show_col(pal, labels = FALSE)
# Continuous, "viridis" from "viridis"
pal <- paletteer_c("viridis::viridis",16)
show_col(pal, labels = FALSE)
# Continuous, "ocean.matter" from "pals"
pal <- paletteer_c("pals::ocean.matter",16)
show_col(pal, labels = FALSE)
# Continuous, "ocean.tempo" from "pals"
pal <- paletteer_c("pals::ocean.tempo",16)
show_col(pal, labels = FALSE)
```

### Diverging

```{r paletteer-div, fig.width=7.5, out.width="60%"}
# Discrete, "TealRose" from "rcartocolor
pal <- paletteer_d("rcartocolor::TealRose")
show_col(pal, labels = FALSE)
# Discrete, "Fall" from "rcartocolor
pal <- paletteer_d("rcartocolor::Fall")
show_col(pal, labels = FALSE)
# Discrete, "RdBu" from "RColorBrewer
pal <- paletteer_d("RColorBrewer::RdBu")
show_col(pal, labels = FALSE)
# Discrete, "BrBG" from "RColorBrewer
pal <- paletteer_d("RColorBrewer::BrBG")
show_col(pal, labels = FALSE)
# Continuous, "vik" from "scico"
pal <- paletteer_c("scico::vik",16)
show_col(pal, labels = FALSE)
# Continuous, "broc" from "scico"
pal <- paletteer_c("scico::broc",16)
show_col(pal, labels = FALSE)
# Continuous, "ocean.balance" from "pals"
pal <- paletteer_c("pals::ocean.balance",16)
show_col(pal, labels = FALSE)
# Continuous, "ocean.curl" from "pals"
pal <- paletteer_c("pals::ocean.curl",16)
show_col(pal, labels = FALSE)
```
