---
title: "Scaling and Saving"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Scaling and Saving}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.asp = 2/3,
  out.width = "75%",
  dev = "ragg_png"
)
```

```{r setup, include = FALSE}
library(ggplot2)
library(crthemes)
```

*Updated `r format(Sys.time(), "%b %d, %Y")`*
\
\
The default `theme_cr()` is designed for a 6" by 4" figure with font size of 12 pt. The `plot_scale` and `font_scale` arguments of `theme_cr()` can be used to change the scaling of the plot features and the font for larger or smaller outputs. 

Note that none of the code in this vignette is run, and the plots are not displayed. If it were run, the saved plot output would be too large to render within a vignette. 

## `plot_scale`

`plot_scale` changes the size of lines, rectangles, `geoms`, and font and the spacing of plot features. It has a default value of 1, which corresponds to a 6" by 4" figure. Larger values will increase the size of the plot features multiplicatively, and smaller values will decrease the size of plot features in the same way. 

The plot size is set with `crsave()`, as described below. 

```{r, eval=FALSE}
# p represents a ggplot object

# Default size
p + theme_cr(plot_scale = 1)

# Larger plot
p + theme_cr(plot_scale = 1.5)
```

## `font_scale`

`font_scale` determines how the font should be sized relative to the rest of the plot. With the default value of 1, the font will be scaled according to the `plot_scale` value. For example, a `plot_scale` value of 2 and a `font_scale` value of 1 will give a font size of 24, corresponding to the increased size of the plot from `plot_scale`.

Changing the `font_scale` value will make the font larger or smaller compared to the plot features. This can be helpful for making font bigger on a small figure where legibility is more important than the size of the data on the figure. 

```{r, eval=FALSE}
# Smaller plot with font twice as large as the default for that plot size
p + theme_cr(plot_scale = 0.5, font_scale = 2)

# Normal plot size with smaller font
p + theme_cr(plot_scale = 1, font_scale = 0.75)
```

## `crsave()`

The `base_scale` and `font_scale` arguments of `theme_cr()` are designed to be used alongside the `crsave()` function. `crsave()` is a wrapper for [`ggsave()`](https://ggplot2.tidyverse.org/reference/ggsave.html) with my preferred graphics devices specified as defaults. 

### Figure Scaling with `base_scale`

The `base_scale` argument corresponds to the `base_scale` argument of the `theme_cr()` function, and it will determine the size of the saved figure unless a value is provided for the `width` argument. The theme is designed for the two arguments to be set to the same value, so the plot features scale appropriately with the size of the plot.

The table below gives some common scaling factors and the figure size in inches they correspond to. 

| Scale Factor | Plot Width | Plot Height |
|:------------:|:----------:|:-----------:|
| 0.5          | 3          | 2           |
| 0.75         | 4.5        | 3           |
| 1            | 6          | 4           |
| 1.5          | 9          | 6           |
| 2            | 12         | 8           |

Note that it is not required that the plot size correspond with `base_scale` value used in `theme_cr()`. For example, for larger plot features, you can use a larger `base_scale` value in `theme_cr()` without changing the save of the saved plot in `crsave()`. This will increase the size of the plot features relative to the figure size. 

The aspect ratio of the plot is set with the `ratio` argument, given as the ratio of the figure width to the figure height. The height of the plot is calculated from the `base_scale` value and the `ratio`.

```{r, eval=FALSE}
# Save the plot object in p with the default settings, defaults to PNG
crsave(p)

# Scale up the plot to a 12" by 8" figure
crsave(p, path = "figure.png", base_scale = 2)
```

### Figure Scaling with `width`

Alternatively, the plot size can be specified directly with the `width` and `units` arguments. In this case, the figure uses the aspect ratio given in `ratio` and the `width` value to set the dimensions of the plot. The `units` argument can be given any unit supported by [`grid::unit()`](https://www.rdocumentation.org/packages/grid/versions/3.6.2/topics/unit).

```{r, eval=FALSE}
# Save the plot as a 4" by 4" figure
crsave(p, path = "figure.png", width = 4, ratio = 1)

# Save the plot as a 5 cm by 10 cm figure
crsave(p, path = "figure.png", width = 5, ratio = 0.5, units = "cm")
```

### Other Arguments

`crsave()` will name the figure file based on the `path` argument or based on the name of the `plot_object` if no `path` is given. It will set the graphics device based on the `device` argument, or it will use the file extension from `path` or the `type` argument to select a default graphics device (recommended). 

`dpi` sets the resolution of the plot in pixels per inch and is passed directly to `ggsave`. Any other arguments for `ggsave` or the specific graphics device being used can also be provided to `crsave()`, and they will be passed through to the relevant function. 

```{r, eval=FALSE}
# Save p as "figure.png"
crsave(p, path = "figure.png")

# Save p as a TIFF file
crsave(p, path = "figure.tiff")

# Decrease the resolution for display on screens
crsave(p, path = "figure.jpg", dpi = 72)
```

### Font Considerations

For `png`, `jpg`, and `tiff` figures, the font is handled through `ragg` and `systemfonts`, and it should work correctly when these packages are installed. See the specifics on the [`ragg` package](https://www.tidyverse.org/blog/2021/02/modern-text-features/) for more information.

For `pdf` and `eps` figures, a cairo-safe font must be used instead of the default theme fonts. This can be accomplished with `theme_cr(cairo = TRUE, font = safefont)` where `safefont` is the name of a system font that work with cairographics (which fonts will work is system-dependent).

For `svg` figures, the font will need to be available on the viewer's computer for it to render properly. There are ways to embed web fonts in the figure itself, but that will have to be handled outside the `crsave()` function. For more information, see the specifics on the [`svglite` package](https://www.tidyverse.org/blog/2021/02/svglite-2-0-0/).

```{r, eval=FALSE}
# Use a cairo-safe font for pdf and eps output
p + theme_cr(cairo = TRUE, font = "Helvetica")
crsave(p, path = "figure.pdf")
```
