---
title: "Scaling and Saving"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Scaling and Saving}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.asp = 2/3,
  out.width = "75%",
  dev = "ragg_png"
)
```

```{r setup, include = FALSE}
library(ggplot2)
library(crthemes)
```

*Updated `r format(Sys.time(), "%b %d, %Y")`*
\
\
The default `theme_cr()` is designed for a 6" by 4" figure with font size of 12 pt. The `plot_scale` and `font_scale` arguments of `theme_cr()` can be used to change the scaling of the plot features and the font for larger or smaller outputs. 

Note that none of the code in this vignette is run, and the plots are not displayed. If it were run, the saved plot output would be too large to render within a vignette. 

## `plot_scale`

`plot_scale` changes the size of lines, rectangles, `geoms`, and font and the spacing of plot features. It has a default value of 1, which corresponds to a 6" by 4" figure. Larger values will increase the size of the plot features multiplicatively, and smaller values will decrease the size of plot features in the same way. 

The plot size is set with `crsave()`, as described below. 

```{r, eval=FALSE}
# p represents a ggplot object

# Default size
p + theme_cr(plot_scale = 1)

# Larger plot
p + theme_cr(plot_scale = 1.5)
```

## `font_scale`

`font_scale` determines how the font should be sized relative to the rest of the plot. With the default value of 1, the font will be scaled according to the `plot_scale` value. For example, a `plot_scale` value of 2 and a `font_scale` value of 1 will give a font size of 24, corresponding to the increased size of the plot from `plot_scale`.

Changing the `font_scale` value will make the font larger or smaller compared to the plot features. This can be helpful for making font bigger on a small figure where legibility is more important than the size of the data on the figure. 

```{r, eval=FALSE}
# Smaller plot with font twice as large as the default for that plot size
p + theme_cr(plot_scale = 0.5, font_scale = 2)

# Normal plot size with smaller font
p + theme_cr(plot_scale = 1, font_scale = 0.75)
```

## `crsave()`

The `plot_scale` and `font_scale` arguments of `theme_cr()` are designed to be used alongside the `crsave()` function to finesse the size of plot components and the final rendered image size. `crsave()` is a wrapper for [`ggsave()`](https://ggplot2.tidyverse.org/reference/ggsave.html) with my preferred graphics devices specified as defaults. 

### The `scaling` argument

Saving rendered images at the desired size and resolution is a tricky business in T. [This `tidyverse` blog post](https://www.tidyverse.org/blog/2020/08/taking-control-of-plot-scaling/) goes into more detail about the different approaches to plot scaling and the pros and cons of each approach. It also explains the methodology behind the `scaling` argument in the `ragg` graphics devices from the [`ragg` package](https://ragg.r-lib.org/). 

The `crsave()` function accepts a `scaling` argument that it then passes to the `ragg` graphics devices if they are being used (the default behavior). The size of every printed component in the image (e.g., points, lines, text, etc.) is essentially then multiplied by this scaling factor to scale up or down the image. 

The `scaling` argument allows the image to either (1) be resized up or down without changing the relative sizes of the individual plot components, or (2) scale up or down the components of the figure relative to the image size.

To resize the image without changing the relative sizes of the plot components, set the scaling factor to the amount the image should be scaled by, and multiply the width of the image by that same factor. For example, to double the size of a 6" x 4" figure without changing the general appearance of the data, set `scaling = 2` and `width = 12`. 

To scale the image components relative to the image size, set the scaling factor to the amount the components should be scaled by without changing the size of the image width. `scaling` > 1 will increase the size of the figure components, and `scaling` < 1 will decrease their size. 

```{r, eval=FALSE}
# Save the plot object in p with the default settings, defaults to PNG
crsave(p)

# Scale up the plot to a 12" by 8" figure
crsave(p, path = "figure.png", width = 12, scaling = 2)
```

### Overall Figure Sizing and Scaling

The most straightforward approach to create figures that are scaled correctly is to first create the figure in R with the `plot_scale` argument of `theme_cr()` set to `1`. With the initial figure creation, determine if any individual aspects of the plot need to be resized relative to the other plot components, and adjust those aspects with the `geom_*()`, `scale_*()`, and `theme()` functions as needed. If the size of the text overall needs adjustment, use the `font_scale` argument of `theme_cr()`. 

After figure is created and the individual components are sized correctly relative to each other in R, the figure can be exported with the `crsave()` function. The next step is to determine the width and aspect ratio for the figure that appears best with the size of the figure components. The approach here is to iteratively tweak the `width` and `ratio` arguments of `crsave()` until the desired figure appearance is reached. 

If the figure is being used alongside previously generated figures and needs to match the relative size of the components in those images, the `plot_scale()` argument of `theme_cr()` can be used to adjust the size of most of the plot components accordingly. This is easiest to use for small adjustments; for any substantial changes in the image scale, use the `scaling` argument of `crsave()`.

The `scaling` argument of `crsave()` is primarily for resizing an existing figure that has already been formatted to the desired appearance. It applies a scaling factor to the entire image, so the `scaling` argument can be used along with the `width` argument to quickly upscale a figure from a presentation to be the correct size for a poster or downscale a figure to make a thumbnail for a website. 

### Image Width and Aspect Ratio

The aspect ratio of the plot is set with the `ratio` argument, given as the ratio of the figure width to the figure height. The height of the plot is calculated from the `width` value and the `ratio`. The `units` argument determines what units are used for the image height and width and can be given any unit supported by [`grid::unit()`](https://www.rdocumentation.org/packages/grid/versions/3.6.2/topics/unit).

```{r, eval=FALSE}
# Save the plot as a 4" by 4" figure
crsave(p, path = "figure.png", width = 4, ratio = 1)

# Save the plot as a 5 cm by 10 cm figure
crsave(p, path = "figure.png", width = 5, ratio = 0.5, units = "cm")
```

### Other Arguments

`crsave()` will name the figure file based on the `path` argument or based on the name of the `plot_object` if no `path` is given. It will set the graphics device based on the `device` argument, or it will use the file extension from `path` or the `type` argument to select a default graphics device (recommended). 

`dpi` sets the resolution of the plot in pixels per inch and is passed directly to `ggsave`. Any other arguments for `ggsave` or the specific graphics device being used can also be provided to `crsave()`, and they will be passed through to the relevant function. 

```{r, eval=FALSE}
# Save p as "figure.png"
crsave(p, path = "figure.png")

# Save p as a TIFF file
crsave(p, path = "figure.tiff")

# Decrease the resolution for display on screens
crsave(p, path = "figure.jpg", dpi = 72)
```

### Font Considerations

For `png`, `jpg`, and `tiff` figures, the font is handled through `ragg` and `systemfonts`, and it should work correctly when these packages are installed. See the specifics on the [`ragg` package](https://www.tidyverse.org/blog/2021/02/modern-text-features/) for more information.

For `pdf` and `eps` figures, a cairo-safe font must be used instead of the default theme fonts. This can be accomplished with `theme_cr(cairo = TRUE, font = safefont)` where `safefont` is the name of a system font that work with cairographics (which fonts will work is system-dependent).

For `svg` figures, the font will need to be available on the viewer's computer for it to render properly. There are ways to embed web fonts in the figure itself, but that will have to be handled outside the `crsave()` function. For more information, see the specifics on the [`svglite` package](https://www.tidyverse.org/blog/2021/02/svglite-2-0-0/).

```{r, eval=FALSE}
# Use a cairo-safe font for pdf and eps output
p + theme_cr(cairo = TRUE, font = "Helvetica")
crsave(p, path = "figure.pdf")
```
